---
title: "Day 1 Recap"
author: "Liana Hardy"
date: '2020-04-22'
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

1. shared / cloud coding environment - good for learning but not good for personal data
2. source control (Github) - track changes and protect origional code. Trace back errors and step back to re-start code. 
3. Reproducible research via literate programming 
4. R
  + packages
  + functions
  + data types
  + objects 
  + tidyverse paradigm
  + logic constructs
  
## Filter exercises 
1. Write a filter that gets all action movies from the movies dataset via the ggplot2movies package
1. Write a filter that removes films lasting more than 6 hours from the movies dataset
1. [ADVANCED] Write a filter that checks to see if any of the films don’t have any genres flagged at all

```{r}
library(tidyverse)
library(ggplot2movies)

ggplot2movies::movies # Can shorten to movies
# Always take a look at the data

```

```{r}
## filter all action movies
# action column is an integer of 0 or 1
movies %>% 
  filter(Action == 1)

```

```{r}
# remove films longer than 6 hours - what is 6 hours in minutes - EXCLUDE
movies %>% 
  filter(length <=6*60) # 6 hours in minutes
  # to check max length - summarise(max(length))
```

```{r}
# no genres flagged
movies %>% 
  filter_at(vars(Action:Short), all_vars( . == 0)) ->
  movies_nogenre

movies_nogenre # here you have assigned your selection to this name using -> 

```

# You got 3/3 GOOD JOB!!!! Check datapipelines.Rmd for answers

## Filtering or selecting columns
```{r}
iris %>% # include these columns
  select(Sepal.Length, Species)

iris %>% # exclude these columns
  select(-Sepal.Length, - Species)

iris %>% # select multiple columns
  select(Sepal.Length:Petal.Width)

iris %>%  # exclude multiple columns
  select(-(Sepal.Length:Petal.Width)) # When in doubt add brackets!!

iris %>% # select depending on start/end/contains
  select(starts_with("Sepal"))

iris %>% # select if a number - 
  select_if(is.numeric)
         
```

## Select Exercises - Section 21
# 1. Output the movies dataset with the column budget changed to budget_if_known
# 1. [ADVANCED] Write a query that returns from the movies dataset columns that have a high degree of missing data (more than 25% of rows are NA) and upper case all the output column names

```{r}
# output movies with renamed column
movies %>% 
  rename(budget_if_known = budget) ->
  movies_cleaned
View(movies_cleaned)

movies %>% 
  select(budger_if_known = budget, everything())

```

## Longer pipelines
Show sepal details for setosa flowers
```{r}
iris %>% 
  filter(Species == "setosa") %>% # Make sure criteria and reduction of information in correct order
  select(starts_with("Sepal"))

```

## Changing or adding columns - using mutate

```{r}

iris %>% 
  mutate(Sepal.Length = Sepal.Length - 0.1, 
         Sepal.Area = Sepal.Length * Sepal.Width,
         Species_trun = str_trunc(as.character(Species),3),
         Diff.Sepal.Length = (Sepal.Length - 
mean(Sepal.Length))/mean(Sepal.Length))
         
# allows change of data but Iris still holds incorrectly 

```

```{r}
# Add a column to movie sthat says how much length differs from the median
movies %>% 
  mutate(diff_length = length - median(length)) ->
  movies_diff
```

## Aggregating our data
- groups = group_by
- aggregation = summarise

```{r}
iris %>% 
  group_by(Species) %>% 
  summarise(count = n(), avg.Sepal.Length = mean(Sepal.Length))

iris %>% 
  group_by(Species) %>% 
  summarise_all(mean)

# Difference in a specific group
iris %>% 
  mutate(Diff.Sepal.Length = (Sepal.Length - mean(Sepal.Length))/mean(Sepal.Length)) %>% 
  group_by(Species) %>% 
  mutate(Species.Diff.Sepal.Length = (Sepal.Length - mean(Sepal.Length))/mean(Sepal.Length))
```

## Find the min and max values for length in movies
## In a pipeline, find out what proportion of films are action, action-comedies and neither action nor comedies
```{r}
# movie length range
movies %>% 
  summarise(min(length), max(length)/60) # divide by 60 to get hours

# proprotions by genre
movies %>% 
  group_by(Action, Comedy) %>% 
  summarise(n=n()) %>% # summarise number of each
  ungroup() %>% # check figures always, use ungroup to stop considering data as group so that mutate works on whole dataset (totals = 100)
  mutate(per=n/sum(n)) # make a percentage
#summarise(sum(per))

```

